---
title: "STA130H1S -- Winter 2021"
author: "N. Moon and S. Caetano [ADD YOUR NAME HERE]"
subtitle: Week 2 Problem Set
output:
  pdf_document: default
---

```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(openintro)
```


# Instructions

## How do I hand in these problems for the January 21st deadline ?

Your complete .Rmd file that you create for these practice problems AND the resulting pdf (i.e., the one you 'Knit to PDF' from your .Rmd file) must be uploaded into a Quercus assignment (link: [https://q.utoronto.ca/courses/206597/assignments/533381](https://q.utoronto.ca/courses/206597/assignments/533381)) by 11:59AM ET, on Thursday, January 21st. Late problem sets are not accepted.



# Problem set grading

There are two parts to your problem set. One is largely R-based with short written answers and the other is more focused on writing. We recommend you use a word processing software like Microsoft Word to check for grammar errors in your written work. Note: there can be issues copying from Word to R Markdown so it may be easier to write in this file first and then copy the text to Word. Then you can make any changes flagged in Word directly in this file.


# Part 1


## [Question 1] The `artwork500.csv` file contains data for a sample of 500 pieces of art owned by the Tate Art Museum. While more variables are available on the Tate Art Museum's site  (github.com/tategallery/collection), you will only be working with the variables featured in the `artwork500.csv` file:
 - id: Unique ID for each piece of artwork
 - artist: Name of the artist
 - title: Title of the artwork
 - type: Medium used
 - year: Year the artwork was created
 - width: width of the artwork, in mm
 - height: height of the artwork, in mm
 - units: measurement units for width and height of the artwork
 - area: surface area (in squared cm)

```{r, message=FALSE}
library(tidyverse) # Load the tidyverse package so it is available to use 
artwork500 <- read_csv("artwork500.csv")
```

### (a) Use the `glimpse()` function to view properties of the `artwork500` data set. How many observations does it include? How many variables are measured for each observation? How many rows and columns does the `artwork500` data frame have?

```{r}

```

*<Type your answer to the written question here (note that R code does not run outside of the grey R code chunks).>*



### (b) Create 3 histograms to explore the distribution of years of creation for this sample of pieces of art: (i) one with 3 bins, (ii) one with 20 bins, and (iii) one with 75 bins; make sure to specify meaningful axis labels where appropriate. Which of these histograms is most appropriate to describe the distribution of the artworks' years of creation? Why? Write a few sentences describing the distribution based on the histogram you chose as most appropriate.

```{r}

```



### (c) Construct a plot to visualize the distribution of a categorical variable and describe the distribution in 1-2 sentences; make sure to specify meaningful axis labels where appropriate. Hint: If you choose a categorical variable with many different categories, you may find it useful to use coord_flip() to flip the bars horizontally and/or change the options in the R code chunk to make the plot large (ex: {r, fig.height=15, fig.width=5}).


```{r}

```


### (d) Construct a set of three boxplots showing visual summaries of the distribution of surface area (`area`) for each type of artwork (`type`); make sure to specify meaningful axis labels where appropriate. Write 3-4 sentences comparing these distributions.

```{r}

```






\newpage

## [Question 2] The `ncbirths` data set is part of the `openintro` package. It consists of observations for a sample of 1000 births in North Carolina in 2004.   Type  `?ncbirths` in the R console for more information about the data and to see the definition of each variable. The code below loads the required libraries for this question and provides a glimpse of the `ncbirths` data frame.
  
```{r, include=FALSE}
library(tidyverse)
library(openintro)
```

```{r}
glimpse(ncbirths)
```
  
  
### (a) Choose two categorical variables and plot their distributions. Identify whether each of these variables is a nominal or ordinal categorical variable. Write one or two sentences interpreting each plot.  

```{r, warning=FALSE, message=FALSE, fig.height=3, fig.width=4, fig.align='center'}

```



### (b) Choose a quantitative variable and plot its distribution.  Identify whether the variable you selected is continuous or discrete, and write 2-3 sentences describing the distribution.

```{r, warning=FALSE, message=FALSE, fig.width=3, fig.height=2, fig.align='center'}

```


### (c) Construct a plot that shows the relationship between birth weight (`weight`) and mother's smoking status (`habit`); make sure to specify meaningful axis labels where appropriate.  

```{r}

```



\newpage

# Part 2

## Writing prompt:  

Pretend that you are on the phone with your friend, and you want to share some of the cool new data visualization techniques that you have been learning in STA130. Pick one of the above graphs and prepare a small paragraph on how you would describe the graph to your friend (keeping in mind that they cannot see the graph). Make sure you include at least 2 words/phrases from your vocabulary list. It is also important to keep in mind that the person you are talking to has not taken STA130, therefore they will not be as familiar with the statistical vocabulary as you are. Therefore, make sure the explain any terms you use in plain language.  

### When describing a figure, it is important to:  

- Describe the data source 
- State the type of graph 
- Identify what are on the x- and y-axis (if appropriate) 
- Describe the distribution 
- Make note of potential outliers  


### Vocabulary List 

- Where are the data centered (approximate values if available) 
- How much spread (relative to what?) 
- Shape: symmetric, left-skewed, right-skewed 
- The tails of the distribution (heavy-tailed or thin-tailed) 
- Modes: where, how many, unimodal, bimodal, multimodal, uniform 
- Outliers, extreme values 
- Frequency (which category occurred the most or least often; data concentrated near a particular value or category) 
- Mean, median, mode 
- Standard deviation, interquartile range 



### Some general reminders 

- Try to not spend more than 20 minutes on the prompt. 
- Aim for more than 200 words but less than 350 words. 
- Use full sentences.
- Grammar is *not* the main focus of the assessment, but it is important that you communicate in a clear and professional manner. I.e., no slang or emojis should appear.
- Be specific. A good principle when responding to a writing prompt in STA130 is to assume that your audience is not aware of the subject matter.
- Remember to end with a conclusion. This means reiterating the key points from your writing sample.

